Carrier phase and amplitude estimation for phase shift keying using pilots and data

### 2012-10-11 paper.tex
minor mods.  This needs work, I need to read Barry's unimodal proof.

We will have particular use of complex random variables of the form $1 + W$ where $W$ is circularly symmetric.  Let $R \geq 0$ and $\Phi \in [0,2\pi)$ be real random variables satisfying, 
\[
R e^{j\Phi} = 1 + W.
\]
The joint pdf of $R$ and $\Phi$ can be shown to be
\begin{equation}\label{eq:pdfRPhi}
f(r,\phi) = \frac{r f_Z(\sqrt{\abs{r^2 - 2r\cos\phi + 1}})}{2\pi\sqrt{\abs{r^2 - 2r\cos\phi + 1}}}.
\end{equation}
The mean of $R e^{j\Phi}$ is equal to one because the mean of $W$ is zero.  So,
\begin{equation}\label{eq:expectRepartRphi}
\expect \Re(R e^{j\Phi}) = \expect R \cos(\Phi) = 1,
\end{equation}
where $\Re(\cdot)$ denotes the real part, and
\begin{equation}\label{eq:expectImpartRphi}
\expect \Im(R e^{j\Phi}) = \expect R \sin(\Phi) = 0,
\end{equation}
where $\Im(\cdot)$ denotes the imaginary part.

@@

\begin{theorem}\label{thm:normality} (Asymptotic normality)
Under the same conditions as Theorem~\ref{thm:consistency}, let $f(r,\phi)$ be the joint probability density function of $R_1$ and $\Phi_1$, let
\[
g(\phi) = \int_{0}^{\infty} r f(r,\phi) dr,
\]
and assume that $\frac{\abs{P}}{L} = p + o(L^{-1/2})$ and $\frac{\abs{D}}{L} = d + o(L^{-1/2})$.
Put $\hat{\lambda}_L = \sfracpart{\theta_0 - \hat{\theta}}_\pi$ and $\hat{m}_L = \hat{\rho} - \rho_0 G(0)$. %If $g(\phi)$ is continuous at $\phi = \tfrac{2\pi}{M}k+\tfrac{\pi}{M}$ for each $k = 0, 1, \dots M-1$, then 
Then the distribution of $(\sqrt{L}\hat{\lambda}_L, \sqrt{L}\hat{m}_L)$ converges to the bivariate normal with zero mean and covariance
\[
\left( \begin{array}{cc} 
\frac{pA_1 + dA_2}{(p + H d)^2} & 0 \\
0 & \rho_0^2(pB_1 + dB_2)
\end{array} \right)
\]
as $L \rightarrow \infty$, where
\[
H = h_2(0) -  2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^{M-1} g(\tfrac{2\pi}{M}k + \tfrac{\pi}{M}),
\]
\[
A_1 = \expect R_1^2\sin^2(\Phi_1), \qquad A_2 = \expect R_1^2\sin^2\fracpart{\Phi_1},
\]
\[
B_1 = \expect R_1^2 \cos^2(\Phi_1) - 1, \;\;\; B_2 = \expect R_1^2 \cos^2\fracpart{\Phi_1} - h_2^2(0).
\]
\end{theorem}

@@

For Theorem~\ref{thm:normality} the integral defining $g(\phi)$ must exist at $\phi = \tfrac{2\pi}{M}k + \frac{\pi}{M}$ for $k = 0, \dots, M-1$ so that the constant $H$ exists.  The next proposition shows that the assumptions made in Theorem~\ref{thm:normality} are strong enough to ensure that the integral does exist at these points.  At the same time we derive some properties that will be useful during the proof of Theorem~\ref{thm:normality}.  The proof is given in Appendix~\ref{appendix:prop1proof}.

\begin{proposition}\label{prop:gg2cont}
Let $W$ be a circularly symmetric complex random variable with $\expect\abs{W}^2$ finite.  Let $R > 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + W$.  Let $f(r,\phi)$ be the joint pdf of $R$ and $\Phi$.  Then
\[
g(\phi) = \int_{0}^\infty r f(r, \phi) dr,
\]
and
\[
g_2(\phi) = \int_{0}^\infty r^2 f(r, \phi) dr
\]
are even functions on $(-\pi,\pi)$ and are continuous at $\phi = \tfrac{2\pi}{M}k + \frac{\pi}{M}$ for $k = 0, \dots, M-1$.
\end{proposition}

@@

\begin{proposition}\label{prop:contgg2}
Let $W = X + jY$ be a circularly symmetric complex random variable, the real and imaginary parts having pdf $f_{X,Y}(x,y)$ that is nonincreasing with $x^2 + y^2$.  Let $R \geq 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + W$.  Let,
\[
h_1(x) = \expect R \cos(x + \Phi) \;\;\; \text{and} \;\;\; h_2(x) =  \expect R \cos\sfracpart{x + \Phi}.
\]
Then $\abs{h_1(x)}$ is uniquely maximised at $x=0$ over the interval $[-\pi,\pi]$ and $\abs{h_2(x)}$ is uniquely maximised at $x = 0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M}]$.
\end{proposition}

@@

\subsection{Proof of Proposition~\ref{prop:gg2cont}} \label{appendix:prop1proof}

\begin{IEEEproof}
BLERG
\end{IEEEproof}

>>>>

We will have particular use of complex random variables of the form $1 + W$ where $W$ is circularly symmetric.  Let $R \geq 0$ and $\Phi \in [0,2\pi)$ be real random variables satisfying, 
\[
R e^{j\Phi} = 1 + W.
\]
The joint pdf of $R$ and $\Phi$ can be shown to be
\begin{equation}\label{eq:pdfRPhi}
f(r,\phi) = \frac{r f_Z(\sqrt{\abs{r^2 - 2r\cos\phi + 1}})}{2\pi\sqrt{\abs{r^2 - 2r\cos\phi + 1}}}.
\end{equation}
Since $\cos\phi$ has period $2\pi$ and is even on $[-\pi,\pi]$ it follows that $f(r,\phi)$ has period $2\pi$ and is even on $[-\pi,\pi]$ with respect to $\phi$.  The mean of $R e^{j\Phi}$ is equal to one because the mean of $W$ is zero.  So,
\begin{equation}\label{eq:expectRepartRphi}
\expect \Re(R e^{j\Phi}) = \expect R \cos(\Phi) = 1,
\end{equation}
where $\Re(\cdot)$ denotes the real part, and
\begin{equation}\label{eq:expectImpartRphi}
\expect \Im(R e^{j\Phi}) = \expect R \sin(\Phi) = 0,
\end{equation}
where $\Im(\cdot)$ denotes the imaginary part.

@@

\begin{theorem}\label{thm:normality} (Asymptotic normality)
Under the same conditions as Theorem~\ref{thm:consistency}, let $f(r,\phi)$ be the joint probability density function of $R_1$ and $\Phi_1$, let
\[
g(\phi) = \int_{0}^{\infty} r f(r,\phi) dr, \qquad g_2(\phi) = \int_{0}^{\infty} r^2 f(r,\phi) dr,
\]
and assume that $\frac{\abs{P}}{L} = p + o(L^{-1/2})$ and $\frac{\abs{D}}{L} = d + o(L^{-1/2})$.
Put $\hat{\lambda}_L = \sfracpart{\theta_0 - \hat{\theta}}_\pi$ and $\hat{m}_L = \hat{\rho} - \rho_0 G(0)$. %If $g(\phi)$ is continuous at $\phi = \tfrac{2\pi}{M}k+\tfrac{\pi}{M}$ for each $k = 0, 1, \dots M-1$, then 
If the functions $g$ and $g_2$ are continuous at $\tfrac{2\pi}{M}k + \tfrac{\pi}{M}$ for each $k = 0, \dots, M-1$, then the distribution of $(\sqrt{L}\hat{\lambda}_L, \sqrt{L}\hat{m}_L)$ converges to the bivariate normal with zero mean and covariance
\[
\left( \begin{array}{cc} 
\frac{pA_1 + dA_2}{(p + H d)^2} & 0 \\
0 & \rho_0^2(pB_1 + dB_2)
\end{array} \right)
\]
as $L \rightarrow \infty$, where
\[
H = h_2(0) -  2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^{M-1} g(\tfrac{2\pi}{M}k + \tfrac{\pi}{M}),
\]
\[
A_1 = \expect R_1^2\sin^2(\Phi_1), \qquad A_2 = \expect R_1^2\sin^2\fracpart{\Phi_1},
\]
\[
B_1 = \expect R_1^2 \cos^2(\Phi_1) - 1, \;\;\; B_2 = \expect R_1^2 \cos^2\fracpart{\Phi_1} - h_2^2(0).
\]
\end{theorem}

@@

Theorem~\ref{thm:normality} requires functions $g$ and $g_2$ to be continuous at $\tfrac{2\pi}{M}k + \tfrac{\pi}{M}$ for each $k = 0, \dots, M-1$.  This places mild restrictions on the distribution of the noise $w_i$.  For example, the requirements are satisfied if the joint pdf of the real and imaginary parts of $w_i$ is continuous, since in this case $f(r,\phi)$ is continuous.  Because $f(r,\phi)$ has period $2\pi$ and is even on $[-\pi,\pi]$ with respect to $\phi$ it follows that both $g$ and $g_2$ have period $2\pi$ and are even on $[-\pi, \pi]$.

%must exist at $\phi = \tfrac{2\pi}{M}k + \frac{\pi}{M}$ for $k = 0, \dots, M-1$ so that the constant $H$ exists.  The next proposition shows that the assumptions made in Theorem~\ref{thm:normality} are strong enough to ensure that the integral does exist at these points.  At the same time we derive some properties that will be useful during the proof of Theorem~\ref{thm:normality}.  %The proof is given in Appendix~\ref{appendix:prop1proof}.

% \begin{proposition}\label{prop:gg2cont}
% Let $W$ be a circularly symmetric complex random variable with $\expect\abs{W}^2$ finite.  Let $R > 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + W$.  Let $f(r,\phi)$ be the joint pdf of $R$ and $\Phi$.  Then
% \[
% g_1(\phi) = \int_{0}^\infty r f(r, \phi) dr,
% \]
% and
% \[
% g_2(\phi) = \int_{0}^\infty r^2 f(r, \phi) dr
% \]
% are even functions on $(-\pi,\pi)$ and are continuous at $\phi = \tfrac{2\pi}{M}k + \frac{\pi}{M}$ for $k = 0, \dots, M-1$.
% \end{proposition}
% \begin{IEEEproof}
% Let $f_Z(z)$ be the cumulative distribution function of $Z = \abs{W}$.  Now $g_1$ and $g_2$ can be written as
% \[
% g_k(\phi) = \int_{0}^{\infty} \frac{\beta(\phi))^{k+1}}{z} f_Z(z) dz
% \]
% where $k = 1, 2$ and $\beta(\phi) = \cos(\phi) + \sqrt{z^2 - 1 + \cos^2(\phi)}$.  Since $\expect \abs{Z}^2$ is finite and non-zero $\int_0^\infty z^c dF(z)$ is finite for all $c = -1,0,1,2$.
% \end{IEEEproof}

@@

\begin{proposition}\label{prop:contgg2}
Let $W = X + jY$ be a circularly symmetric complex random variable, the real and imaginary parts having continuous and differentiable joint pdf $f_{X,Y}(x,y)$ that is nonincreasing with $x^2 + y^2$.  Let $R \geq 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + W$.  Let,
\[
h_1(x) = \expect R \cos(x + \Phi) \;\;\; \text{and} \;\;\; h_2(x) =  \expect R \cos\sfracpart{x + \Phi}.
\]
Then $\abs{h_1(x)}$ is uniquely maximised at $x=0$ over the interval $[-\pi,\pi]$ and $\abs{h_2(x)}$ is uniquely maximised at $x = 0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M}]$.
\end{proposition}

@@

% \subsection{Proof of Proposition~\ref{prop:gg2cont}} \label{appendix:prop1proof}

% \begin{lemma}
% Let $W$ be a circularly symmetric random variable with constant magnitude $Z = \abs{W} = c > 0$ and phase uniformly distributed on $[-\pi,\pi)$.  Let $R \geq 0$ and $\Phi \in [-\pi, \pi)$ be real random variables such that $Re^{j\Phi} = 1 + W$.  Then the pdf of $\Phi$ is
% \begin{equation} \label{eq:fphicleq1}
% f_{\Phi}(\phi) = \begin{cases} 
% \frac{\cos(\Phi)}{\pi c} &  [-\tfrac{\pi}{2}, \tfrac{\pi}{2}] \\
% 0 & \text{otherwise}
% \end{cases}
% \end{equation}
% when $0 < c \leq 1$, and
% \begin{equation} \label{eq:fphicg1}
% f_{\Phi}(\phi) = \frac{\cos(\phi) + \sqrt{c^2 - 1 + \cos^2(\phi)}}{2\pi c}
% \end{equation}
% when $c > 1$.
% \end{lemma}
% \begin{IEEEproof}
% We make use of delta `function' notation and write the `pdf' of $Z$ as $\delta(z - c)$.  Using~\eqref{eq:pdfRPhi} the pdf of $\Phi$ is
% \[
% f_{\Phi}(\phi) = \int_{0}^{\infty} r \frac{\delta(z - c)}{ 2\pi z} dr = \sum_{r} \frac{r}{2\pi c}
% \]
% where $z^2 = r^2 - 2r\cos(\phi) + 1$ and the sum is over the positive real solutions of $c^2 = r^2 - 2r\cos(\phi) + 1$.  If $0 < c \leq 1$ and $\phi \in [-\tfrac{\pi}{2}, \tfrac{\pi}{2}]$ there are two positive real solutions, these being
% \[
% r = \cos(\phi) \pm \sqrt{c^2 - 1 + \cos^2(\phi)},
% \]
% whilst if $\phi \notin [-\tfrac{\pi}{2}, \tfrac{\pi}{2}]$ there are no real solutions.  Thus~\eqref{eq:fphicleq1} holds.  If $c > 1$ there is a single real solution
% \[
% r = \cos(\phi) + \sqrt{c^2 - 1 + \cos^2(\phi)}
% \]
% so that~\eqref{eq:fphicg1} holds.
% \end{IEEEproof}

% Now if $W$ is a circularly symmetric complex random variable with $Z = \abs{W}$ having pdf $f_Z$ and let $R \geq 0$ and $\Phi \in [-\pi, \pi)$ be real random variables such that $Re^{j\Phi} = 1 + W$.  The pdf of $\Phi$ is 
% \[
% f_{\Phi}(\phi) = \int_{0}^{\infty} r\frac{f_Z(z)}{ 2\pi z} dr.
% \]
% Let
% \[
% I_{N,c}(z) = \begin{cases}
% N \frac{f_Z(z)}{ 2\pi z} & z \in [cN, c(N-1)) \\
% 0 & \text{otherwise}.
% \end{cases}
% \]
% Now
% \[
% \frac{f_Z(z)}{ 2\pi z} = \frac{1}{N} \sum_{c = 0}^{\infty}I_{N,c}(z)
% \]
% so that
% \[
% f_{\Phi}(\phi) = \int_{0}^{\infty} r\frac{1}{N} \sum_{c = 0}^{\infty}I_{N,c}(z) dr = \frac{1}{N} \sum_{c = 0}^{\infty} \int_{0}^{\infty} r I_{N,c}(z) dr 
% \]
